### 2018-03-22 doc/ijcai-18/rebuttal/rebuttal.txt
updates added to IJCAI18 rebuttal

Rebuttal #12637:
-------------
We sincerely thank reviewers for their constructive comments and criticisms which help to clarify details of our approach. This is the first work that LEARNS and EVALUATES action models exclusively using an off-the-shelf classical planner. We belief this result is relevant by itself because it opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.

The SAT-based system ARMS (AIJ07) is the closest system to compare with, since ARMS also requires the operator heads and observations of correct plans. ARMS addresses however a different learning task:
(1) ARMS learns from plan traces (where the ACTIONS observations are assumed to be correct) while we learn from STATE observations.
(2) ARMS addresses an OPTIMIZATION task (i.e. maximizing the # of covered examples) while ours is a SATISFYING task. We compute a conservative model that is guaranteed to be applicable at all the given observations.

Rebuttal #23413:
-------------
In effect, we definitely agree with the reviewer that a SAT-based compilation could also be used for the learning and evaluation tasks addressed in this paper. The only reason why we bet on planning is because, when there is an unknown # of missing observations, the planning horizon is no longer known and addressing this learning task is in our research agenda. Further, the planning community has shown terrific expertise in the development of informative while inexpensive heuristics (current heuristics can evaluate millions of symbolic states in seconds time). Our believe is that this technology has a key role to play in the assessment of generative models with regard to large amounts of incomplete/noisy examples.

Like happens with satisfying planners, our approach is extensible to the corresponding optimization task, e.g. in our case minimizing the # of model edits required to cover a set of observations. As a first step, this paper aimed at understanding the grounds of our compilation for learning and evaluating STRIPS models in fully-observable environments (e.g. classical planning tasks) but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.

Answering the reviewer questions:
1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained slightly better experimental results. There is nothing that prevent us from starting with the most general hypothesis (zero preconditions). In fact the evaluation task starts with a given action model that may have for instance, no preconditions. 
2. We agree with the reviewer that the "one test_i fluent" encoding is more natural. We plan to evaluate it and check whether it allow us to achieve some performance gain.
To comments on the training set we refer the reviewer to the Rebuttal #26170.

Rebuttal #25749:
-------------
The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains (15 by the time being despite we are working to cover any IPC domain with the STRIPS requirement) but also for EVALUATING how well a given STRIPS model matches a test set of observations.

Evaluating a given STRIPS model according to a test set of observations enables the "STRIPS model recognition" task. This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape). Different generative models like policies, programs, grammars and different forms of domain-specific control knowledge are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al). With this regard, our work poses a general framework to assess how a given generative model (provided that it is STRIPS compilable) matches a given set of observations.

Rebuttal #26170:
-------------
Our evaluation focused on understanding which models can be learned from just 25 observations. Results revealed that 25 was not enough to learn action models in the domains with the largest # of schemas. In our evaluation this # ranges [1,7]. The largest this # the lowest chance of choosing an operator, e.g. "disembark-truck" is missing from the driverlog or "paint-down" is missing from floortile. Including more observations, specially of the missing schemes, leads to better models. Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance. Current SATplanners have performance issues with horizons beyond 150-200 steps. Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (we require 2 actions to validate 1 observation, Fig 3).

Precision&Recall did not reach 1.0 in some domains with few schemes because learning discovered preconditions that were not specified in the IPC/reference domain. For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, while only (connected ?place1 ?place2) was specified in the IPC model. Effects were perfectly learned in these domains.

The input observations affect to the quality of the learned models. Were use 1 problem for each domain with a small # of objects because this way the compiled task is more tractable. If IPC instances are small enough we use them otherwise, we created a smaller problem. An open question is identifying the smallest # of objects per type required to learn a complete STRIPS model. This has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, Rintanen IJCAI17). We believe this research direction is promising to obtain similar results for learning STRIPS models.