### 2018-06-04 doc/aaai-19/model-recognition.tex
updates added to aaai19 proposal

\def\year{2017}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{}
\setcounter{secnumdepth}{0}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{comment}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

\lstset{
  basicstyle=\ttfamily,
  mathescape
}

\usepackage{multicol}
\usepackage{arydshln}
\usetikzlibrary{calc,backgrounds,positioning,fit}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\newcommand{\pre}{\mathsf{pre}}     % precondition
\newcommand{\del}{\mathsf{del}}     % effect
\newcommand{\add}{\mathsf{add}}     % effect
\newcommand{\eff}{\mathsf{eff}}     % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\true}{\mathsf{true}}   % true
\newcommand{\false}{\mathsf{false}} % false
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}     % precondition

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

\begin{document}

\title{Model Recognition as Planning}

% Commented for blind submission
\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia \and Miquel Ram\'irez\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\`ecnica de Val\`encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}

\maketitle
\begin{abstract} 
Given a set of planning models and a partially observed plan execution, we define {\em model recognition} as the task of identifying which model in the set has the highest probability of producing the input observation. The paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ model to produce a partially observed plan execution. This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner. The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em Turing Machines}. We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.
\end{abstract}

\section{Introduction}
\label{sec:introduction}
{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behaviour. Plan recognition is considered {\em automated planning} in reverse; while automated planning aims to compute a sequence of actions that accounts for a given goals, plan recognition aims to compute the goals that account for an observed sequence of actions~\cite{geffner:book:2013}.

Diverse approaches has been proposed for plan recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{carberry2001techniques}. {\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}. This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal according to the observed plan execution.

In this paper we introduce the task of {\em model recognition}.  Given a set of planning models and a partially observed plan execution, {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation. {\em Model recognition} is of interest because:
\begin{itemize}
\item Once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
\item It enables recognizing algorithms by observing their execution. Planning models can encode diverse forms of algorithm representation, like {\em {\sc GOLOG} programs}, {\em finite state controllers} or {\em push-down automata}~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
\end{itemize}

This paper introduces {\em model recognition as planning}, a novel method to assess the probability of a given \strips\ model to produce an observed plan execution. The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner. The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em Turing Machines} (all of them defined within the same {\em tape alphabet} and same {\em machine states} but different {\em transition functions}). We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.

\section{Background}
\label{sec:background}
This section formalizes classical planning and characterizes the space of \strips\ models.

\subsection{Classical planning}
We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values). We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.

A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$, so the size of the state space is $2^{|F|}$. Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often we will abuse of notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$. We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$. The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The {\em plan length} is denoted with $|\pi|=n$ . A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.

\subsection{The space of \strips\ models}
Like in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$. Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^k$ is the $k$-th Cartesian power of $\Omega$. We assume also that actions $A$ are instantiated from a set of action schemas. Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
\begin{figure}
\begin{scriptsize}
\begin{verbatim}
(:action stack
 :parameters (?v1 ?v2 - object)
 :precondition (and (holding ?v1) (clear ?v2))
 :effect (and (not (holding ?v1)) (not (clear ?v2))
              (handempty) (clear ?v1) (on ?v1 ?v2)))
\end{verbatim}
\end{scriptsize}
 \caption{\small \strips\ action schema coded in PDDL.}
\label{fig:stack}
\end{figure}

Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be an additional set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame. For instance, in a three-block four-operator {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two. We define $F_v$, a new set of fluents, $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}. In the mentioned {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.

We define an \strips\ action schema as the tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
\begin{itemize}
\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$. The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}

For a given $head(\xi)$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the preconditions and effects of the corresponding action schema. For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element from $F_v$ that involves $v_2$ because {\small\tt pickup($v_1$)} has arity one. The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_v(\xi)|}$. Note that \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition. For the mentioned {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while $2^{2|F_v(pickup)|}=1024$.

We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi')$). For instance, we claim that blocksworld operators {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not. Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.

\section{Model Recognition}
\label{sec:recognition}
Given a finite and non-empty set of {\em comparable} action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, and a partially observed plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates $s_i$. {\em Model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.

\subsection{The \strips\ edit distance}
Our approach for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by $\mathcal{M}$ to produce $\mathcal{T}$.

In this work we assume that action models $\mathcal{M}\in M$ are represented with the \strips\ formalism. We define here the two allowed \emph{operations} to edit an \strips\ action model:
\begin{itemize}
\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_v(\xi)$.
\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_v(\xi)$.
\end{itemize}

We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are. The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.

\begin{definition}
  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models. The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
\end{definition}

Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well. In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
\begin{definition}
The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
\end{definition}

We define now an edit distance to asses the matching of a learned action model with respect to a plan trace $\mathcal{T}$.

\begin{definition}
  Given $\mathcal{M}$, a \strips\ action model built from $F_v$ and $\mathcal{T}$, a plan trace built with fluents in $F$ and actions in $A$. The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t. $\mathcal{M}'$ produces a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
\end{definition}

The $\delta(\mathcal{M},\mathcal{T})$ distance could also be defined assessing the edition required by the observed plan execution to match the given model. This implies defining {\em edit operations} that modify $\mathcal{T}$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}. Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$. The reason is that in practice, the number of {\em variable objects} is normally smaller than the number of objects in the observed states.

\subsection{The \strips\ probability distribution}
According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$. In our scenario, the hypotheses are about the set of possible \strips\ action models $\mathcal{M}\in M$ while the given observation is a partially observed plan execution $\mathcal{T}$.

We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of the comparable \strips\ models (the definable models within the $F_v(\xi)$ sets) provided that the plan execution $\mathcal{T}$ is observed. This probability distribution can be estimated in three-steps:
\begin{enumerate}
\item Estimating the {\em a priori} probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$. For instance, given the set of predicates $\Psi$ and the given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$. If we assume that a priori all models are equiprobable this means that $P(\mathcal{M})=\frac{1}{\prod_\xi 2^{2|F_v(\xi)|}}$. With respect to the observations, given the sets $F$ and $A$, the number of possible state observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|F|}$ while the number of possible action observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|A|}$. If we assume that a priori all traces are equiprobable, this means $P(\mathcal{T})=\frac{1}{2^{2n\times|F|\times |A|}}$.
\item Estimating the conditional probability $P(\mathcal{T}|\mathcal{M})$. We can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and map  this distance into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$. 
\item Applying the Bayes rule to obtain the normalized posterior probabilities since these probabilities must sum 1.
\end{enumerate}

\section{Model Recognition as Planning}
\label{sec:asPlanning}
For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$  using a classical planning compilation. The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}. 
\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}. The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$.
\end{enumerate}

Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing. The edited action model is validated at a four action plan for inverting a two-block tower where states $s_0$ and $s_4$ are observed while the intermediate states $s_1$, $s_2$ and $s_3$ are unobserved. 
\begin{figure}
{\tt\scriptsize
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)\\
02 : (apply\_unstack blockB blockA i1 i2)\\
03 : (apply\_putdown blockB i2 i3)\\
04 : (apply\_pickup blockA i3 i4)\\
05 : (apply\_stack blockA blockB i4 i5)\\
06 : (validate\_1)
}
 \caption{\small Plan for editing (steps [0-1]) and validating (steps [2-6]) a given {\em blockswold} model.}
\label{fig:plan-pdistance}
\end{figure}

Our interest is not in the resulting action model from the compilation $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g. $\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}. In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$. %The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.

\subsection{Conditional effects}
Conditional effects allow us to compactly define the actions output by our compilation. An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$. Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}. An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
\[
triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
\]

The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.

\subsection{The compilation formalization}

Given $\tup{\mathcal{M},\Psi,\mathcal{T}}$ s.t.,
\begin{itemize}
\item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
\item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t. {\small $1\leq i\leq n$}, can be {\em partially observed}. This means that some fluents in $s_i$ may be missing because its value is unknown. In the extreme, entire states $s_i$, {\small $1\leq i\leq n$}, can be missing. 
\item $\mathcal{T}$ is {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
\end{itemize}
the compilation outputs a classical planning task $P=\tup{F,A,I,G}$:
\begin{itemize}
\item $F$ contains:
\begin{itemize}
\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace given as input, i.e. the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:plan-pdistance}. Formally, $\Omega=\bigcup_{\small s\in\mathcal{T}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model. If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$. For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\mathcal{T}$. The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\mathcal{T}$.
\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
\end{itemize}
\item $I$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated. Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect. Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.

\item $G=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in the state observations $s_i\in\mathcal{T}$.
\item $A$ comprises three kinds of actions:
\begin{enumerate}
\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
\begin{itemize}
\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.

\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}, pre_{f}(\xi)\},\\
\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
\end{align*}
\end{small}

\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.

\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
\end{align*}
\end{small}
\end{itemize}
Besides these actions $A$ also contains the actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect.

\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$. Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position. Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
&\cup \{\neg mode_{val}\},\\
\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
&\{\emptyset\}\rhd\{mode_{val}\}.
\end{align*}
\end{small}

\begin{figure}[hbt!]
\begin{scriptsize}
\begin{verbatim}
(:action apply_stack
  :parameters (?o1 - object ?o2 - object)
  :precondition
   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
        (or (not (pre_clear_stack_v1)) (clear ?o1))
        (or (not (pre_clear_stack_v2)) (clear ?o2))
        (or (not (pre_holding_stack_v1)) (holding ?o1))
        (or (not (pre_holding_stack_v2)) (holding ?o2))
        (or (not (pre_handempty_stack)) (handempty)))
  :effect
   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
        (when (del_ontable_stack_v1) (not (ontable ?o1)))
        (when (del_ontable_stack_v2) (not (ontable ?o2)))
        (when (del_clear_stack_v1) (not (clear ?o1)))
        (when (del_clear_stack_v2) (not (clear ?o2)))
        (when (del_holding_stack_v1) (not (holding ?o1)))
        (when (del_holding_stack_v2) (not (holding ?o2)))
        (when (del_handempty_stack) (not (handempty)))
        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
        (when (add_ontable_stack_v1) (ontable ?o1))
        (when (add_ontable_stack_v2) (ontable ?o2))
        (when (add_clear_stack_v1) (clear ?o1))
        (when (add_clear_stack_v2) (clear ?o2))
        (when (add_holding_stack_v1) (holding ?o1))
        (when (add_holding_stack_v2) (holding ?o2))
        (when (add_handempty_stack) (handempty))
        (when (modeProg) (not (modeProg)))))
\end{verbatim}
\end{scriptsize}
 \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
\label{fig:compilation}
\end{figure}
The extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\

\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{T}$, {\tt\small $1\leq i< n$}.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
\end{align*}
\end{small}
\end{enumerate}
\end{itemize}

\subsection{Evaluation}
\label{sec:evaluation}
To evaluate the empirical performance of {\em model recognition as planning} we defined a set of possible \strips\ models, each representing a different {\em Turing Machine}, but all sharing the same set of machine states and same tape alphabet.

\subsection{Modeling {\em Turing Machines} with \strips\ }
A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma,\Upsilon,\square,\delta}$:
\begin{itemize}
\item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ is the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of acceptor states.  
\item $\Sigma$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that contains the {\em input alphabet} $\Upsilon\subseteq\Sigma$ (the subset of symbols allowed to initially appear in the tape) and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}. For each possible pair of tape symbol and machine state, $\delta$ defines (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation. If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
\end{itemize}

Figure~\ref{tab:tm-anbncn} shows the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language. The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.

\begin{figure}
\begin{center}
    \begin{tabular}{| c || c | c | c | c | c | c |}
    \hline
      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline\hline
    a & x,r,$q_1$ & a,r,$q_1$ & - &  a,l,$q_3$ & - & - \\ \hline
    b & - & y,r,$q_2$ & b,r,$q_2$ & b,l,$q_3$ & - & - \\ \hline
    c & - & - & z,l,$q_3$ & - & - & - \\ \hline
    x & - & - & - & x,r,$q_0$ & - & - \\ \hline
    y & y,r,$q_4$ & y,r,$q_1$ & - & y,l,$q_3$ & y,r,$q_4$ & - \\ \hline
    z & - & - & z,r,$q_2$ & z,l,$q_3$ & z,r,$q_4$ & - \\\hline
    $\square$ & - & - & - & - & $\square$,r,$q_5$  & - \\                
    \hline
    \end{tabular}
\end{center}
  \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
  \label{tab:tm-anbncn}
\end{figure}

A classical planning frame $\Phi=\tup{F,A}$ can encode the {\em transition function} $\delta$ of a {\em Turing Machine} $\mathcal{M}$ as follows:
\begin{itemize}
\item Fluents $F$ are instantiated from a set of four {\em predicates} $\Psi$: {\small\tt (head ?x)} that encodes the current position of the header in the tape. {\small\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape. {\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$. {\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state. Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ schema is defined:
\begin{itemize}
\item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$. Parameters $?xl$, $?x$ and $?xr$ are tape cells.
\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl/?xr$ its left/right neighbours. Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol pointed by the header and the currrent machine state.
\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
\end{itemize}
\end{itemize}

The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} (the full encoding of the {\em Turing Machine} of Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema). 
\begin{figure}[hbt!]
\begin{scriptsize}
\begin{lstlisting}
(:action transition-1      ;;; a,$q_0\rightarrow$ x,r,$q_1$
  :parameters (?xl ?x ?xr)
  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
                     (next ?xl ?x) (next ?x ?xr))
  :effect (and (not (head ?x)) 
               (not (symbol-a ?x)) (not (state-$q_0$))
               (head ?xr) (symbol-x ?x) (state-$q_1$)))
\end{lstlisting}
\end{scriptsize}
 \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
\label{fig:update-rule}
\end{figure}

Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of a {\em Turing Machine} are definable as {\em plan traces} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.

\subsection{Experimental setup}
We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a two-symbol three-state {\em Turing Machine}. We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$. Finally, we follow our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\mathcal{T}$. This exeperiment is repeated for different amounts of missing information in the input trace $\mathcal{T}$.

\subsubsection{Reproducibility}
{\sc Madagascar} is the classical planner we used to solve the instances that result from our compilations for its ability to deal with dead-ends~\cite{rintanen2014madagascar}. Due to its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because there is no interaction between these actions. Actions for programming effects can also be applied in a single planning step, thus significantly reducing the planning horizon.

The compilation source code, evaluation scripts and benchmarks (including the used training and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.

\subsection{Results}

\subsection{Conclussions}
\label{sec:conclussions}

The {\em \strips\ probability distribution} assigns the same probabaility value to any hypothesis model $\mathcal{M}\in M$ that can reproduce the given observation $\mathcal{T}$. As a rule of thumb, the less information observed the more different models can reproduce observations. If we assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}, one can define probability distributions that consider as more probable hypothesis the ones that reproduce observations in less steps. %This an interesting research direction, specially for the particular case $\mathcal{T}=\tup{s_0, s_n}$ where observations are {\em input-output} examples.

Recently {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} also proposes model edition but in this case the aim is to conform the different PDDL models of two different agents.

\bibliographystyle{aaai}
\bibliography{planlearnbibliography}

\end{document}