### 2018-06-19 doc/aij18/evaluation.tex
updates added to section 5

\section{Evaluating \strips\ Action Models with Classical Planning}
\label{sec:evaluation}

If the actual GTM model is available then the quality of learned action model is quantifiable using the well-studied ML metrics, {\em precision} and {\em recall}. {\em Precision} and {\em recall} are two syntactic metrics frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative than counting the number of errors between the learned and the GTM model~\cite{davis2006relationship}. Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
\begin{itemize}
\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (in our case, predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing).
\end{itemize}

When learning action models from observations of plan executions the semantic of the learned models can be easy altered. For instance, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped. These role swaps typically happen when the observed input data, given in $\tau$, is scarce (e.g. if $\tau$ does not contain any intermediate action, the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa). Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be semantically correct with respect to the given input observations.

Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM; i.e. a learned model semantically equivalent but syntactically different to the GTM. Here we introduce a novel evaluation approach that is robust to role changes of this particular kind. The intuition of the approach is to {\em semantically} assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations. This semantic evaluation approach is again flexible to various amount and kind of available input knowledge.

\subsection{The \strips\ edit distance}

We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
\begin{itemize}
\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_{\xi}$.
\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_{\xi}$.
\end{itemize}

We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are. The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.

\begin{mydefinition}
  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models. The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
\end{mydefinition}

Note that the number of {\em deletions} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$ matches our previous definition for the number of {\em false positives} when: (1) $\mathcal{M}$ is a learned model and (2), $\mathcal{M}'$ is the given GTM that serves as reference. Likewise the number of insertions required to transform $\mathcal{M}$ into $\mathcal{M}'$ corresponds to the number of {\em false negatives} in the learned model $\mathcal{M}$ with respect to the GTM $\mathcal{M}'$. As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a given learned model with respect to a reference model. Further, we can claim that all these numbers are bounded. Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well. In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_{\xi}|$. This is the same max number of edits that can be introduced to the positive and to the negative effects of $\xi\in\mathcal{M}$.
\begin{mydefinition}
The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is given by the expression $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_{\xi}|$.
\end{mydefinition}

In practice evaluating a given learned domain with respect to the actual GTM model is often impossible because the actual GTM model is unknown. The ARMS system showed that the error of a learned action model can also be estimated with respect to a set of observations of plan executions by assuming that the observations were generated with the learned model~\cite{yang2007learning}. Inspired by this approach for the semantic evaluation of learned action models, we define now an edit distance to asses the quality of a learned action model with respect to a partially observed plan trace.

\begin{mydefinition}
  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and a plan trace $\tau=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{m}}$ whose state observation are built with fluents in $F$ and observed actions are all in $A$. The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan trace $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
\end{mydefinition}

Unlike the semantic metrics defined by ARMS, thes $\delta(\mathcal{M},\tau)$ distance does not accumulate errors that are caused by the same flaw in the learned model $\mathcal{M}'$. Further, our {\bf observation edit distance}, allows us to define a semantic version of the well-studied {\em precision} and {\em recall} metrics. Following the previous correspondences ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can define a {\em semantic} version of the {\em precision} and {\em recall} metrics of a learned model. In this case {\em precision} and {\em recall} is not computed with respect to a GTM but with respect to a partially observed plan trace $\tau=\tup{s_0,a,_1,s_1,\ldots,a_n,s_m}$. To define the {\em semantic} version of {\em precision} and {\em recall} the only term to re-formalize is the number of {\em true positives} that, for each action schema $\xi\in\mathcal{M}$, is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$, where $fp(\xi)$ is the number of deletions done in the schema $\xi\in\mathcal{M}$.

\subsection{Computing edit distances with classical planning}
Our compilation is extensible to compute the {\em observation edit distance} and hence, our semantic versions of the {\em precision} and {\em recall} metrics. This extension considers that the input \strips\ model $\mathcal{M}$, is {\em non-empty} so instead of learning an action model from scratch we simply edit $\mathcal{M}$ until it satisfies the given input observations. In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets. A solution to the classical planning task resulting from the extended compilation is a sequence of actions that:
\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}. The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:learning} for the models that are programmed from scratch.
\end{enumerate}

Given $\Lambda=\tup{\mathcal{M},\tau}$, the output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda},G_{\Lambda}}$:
\begin{itemize}
\item $F_{\Lambda}$, $I_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation. Note that, the input action model $\mathcal{M}$ is encoded in the initial state. This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in $I_{\Lambda}$ iff they appear in $\mathcal{M}$.
\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$. The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation. The only difference here is that each action for {\em programming} an operator schema now implement the two {\em edit operations} (i.e. include also the actions for {\em deleting} a precondition or negative/positive effect from a given action schema).
\end{itemize}

Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where only the two positve effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing. In this case the edited action model is again validated in the plan and state observations shown in Figure~\ref{fig:example-plans}.

\begin{figure}[hbt!]
{\footnotesize\tt
  {\bf 00} : (insert\_add\_handempty\_stack)\\
  01 : (insert\_add\_clear\_stack\_var1)\\
  {\bf 02} : (apply\_unstack blockB blockA i1 i2)\\
  03 : (apply\_putdown blockB i2 i3)\\
  04 : (apply\_pickup blockA i3 i4)\\
  05 : (apply\_stack blockA blockB i4 i5)\\
  {\bf 06} : (validate\_1)
}
\caption{\small Plan for editing and validating a blockswold action model where only the positve effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.}
\label{fig:plan-pdistance}
\end{figure}

Our interest when solving the classical planning task that is output by our extended compilation is not in the resulting action model $\mathcal{M}'$ but in the number of {\em edit operations} (insertions and deletions) that is required to compute $\mathcal{M}'$. For instance, in the example of Figure~\ref{fig:plan-pdistance}, $\mathcal{M}'$ is validated in the given observations after inserting two new positive effects to the initial action model, e.g. $\delta(\mathcal{M},\mathcal{T})=2$ . In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$  (as shown in Section~\ref{sec:learning}). The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.

Last but not least, the compilation approach is also flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$. A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion. In this case the edited model is not validated on a sequence of observations or plans but on the given action model $\mathcal{M}'$ that acts as a reference. The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation. With respect to the actions, $A_{\Lambda}'$ again implement the two {\em edit operations} ({\em insertion} and {\em deletion} for both preconditions and effects) but does not contain $\mathsf{apply_{\xi,\omega}}$ or $\mathsf{validate_{i}}$ actions because the edited \strips\ action model are not validated in any observation of plan executions. Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_{\xi}$. To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model. The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects. Again our interest is in the number of required {\em edit operations}, e.g. $\delta(\mathcal{M},\mathcal{M'})=2$.

\begin{figure}[hbt!]
{\tt\small
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)
}
 \caption{\small Plan for computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.}
\label{fig:plan-mdistance}
\end{figure}