### 2018-12-04 doc/aij18/major-revision/appendix.tex

\section*{Appendix}
\label{sec:appendix}

\begin{figure}[hbtp!]
\begin{scriptsize}  
\begin{verbatim}
(define (domain BLOCKS)
  (:requirements :strips)
  (:predicates (on ?x ?y)(ontable ?x)(clear ?x)(handempty)(holding ?x))

  (:action pick-up
	     :parameters (?x)
	     :precondition (and (clear ?x) (ontable ?x) (handempty))
	     :effect (and (not (ontable ?x)) (not (clear ?x)) (not (handempty)) (holding ?x)))

  (:action put-down
	     :parameters (?x)
	     :precondition (holding ?x)
	     :effect (and (not (holding ?x)) (clear ?x) (handempty) (ontable ?x)))

  (:action stack
	     :parameters (?x ?y)
	     :precondition (and (holding ?x) (clear ?y))
	     :effect (and (not (holding ?x)) (not (clear ?y)) (clear ?x) (handempty) (on ?x ?y)))

  (:action unstack
	     :parameters (?x ?y)
	     :precondition (and (on ?x ?y) (clear ?x) (handempty))
	     :effect (and (holding ?x) (clear ?y) (not (clear ?x)) (not (handempty)) (not (on ?x ?y)))))
  \end{verbatim}
\end{scriptsize}  
\caption{\small PDDL domain file for the blocksworld domain.}
\label{fig:bw-domain}
\end{figure}

Compiled PDDL domain file for learning the blocksworld action models from two initial and final states.

\begin{scriptsize}  
\begin{verbatim}
(define (domain blocks)
 (:requirements :strips)
 (:types object - None )
 (:constants a -  object b -  object c -  object d -  object e -  object f -  object g -  object)
 (:predicates (add_clear_pick-up_var1) (add_clear_put-down_var1) (add_clear_stack_var1) (add_clear_stack_var2) 
(add_clear_unstack_var1) (add_clear_unstack_var2) (add_handempty_pick-up) (add_handempty_put-down) 
(add_handempty_stack) (add_handempty_unstack) (add_holding_pick-up_var1) (add_holding_put-down_var1) 
(add_holding_stack_var1) (add_holding_stack_var2) (add_holding_unstack_var1) (add_holding_unstack_var2) 
(add_on_stack_var1_var1) (add_on_stack_var1_var2) (add_on_stack_var2_var1) (add_on_stack_var2_var2) 
(add_on_unstack_var1_var1) (add_on_unstack_var1_var2) (add_on_unstack_var2_var1) (add_on_unstack_var2_var2) 
(add_ontable_pick-up_var1) (add_ontable_put-down_var1) (add_ontable_stack_var1) (add_ontable_stack_var2) 
(add_ontable_unstack_var1) (add_ontable_unstack_var2) (clear ?o1 - object) (del_clear_pick-up_var1) 
(del_clear_put-down_var1) (del_clear_stack_var1) (del_clear_stack_var2) (del_clear_unstack_var1) 
(del_clear_unstack_var2) (del_handempty_pick-up) (del_handempty_put-down) (del_handempty_stack) 
(del_handempty_unstack) (del_holding_pick-up_var1) (del_holding_put-down_var1) (del_holding_stack_var1) 
(del_holding_stack_var2) (del_holding_unstack_var1) (del_holding_unstack_var2) (del_on_stack_var1_var1) 
(del_on_stack_var1_var2) (del_on_stack_var2_var1) (del_on_stack_var2_var2) (del_on_unstack_var1_var1) 
(del_on_unstack_var1_var2) (del_on_unstack_var2_var1) (del_on_unstack_var2_var2) (del_ontable_pick-up_var1) 
(del_ontable_put-down_var1) (del_ontable_stack_var1) (del_ontable_stack_var2) (del_ontable_unstack_var1) 
(del_ontable_unstack_var2) (handempty) (holding ?o1 - object) (modeProg) (on ?o1 - object ?o2 - object) 
(ontable ?o1 - object) (pre_clear_pick-up_var1) (pre_clear_put-down_var1) (pre_clear_stack_var1) 
(pre_clear_stack_var2) (pre_clear_unstack_var1) (pre_clear_unstack_var2) (pre_handempty_pick-up) 
(pre_handempty_put-down) (pre_handempty_stack) (pre_handempty_unstack) (pre_holding_pick-up_var1) 
(pre_holding_put-down_var1) (pre_holding_stack_var1) (pre_holding_stack_var2) (pre_holding_unstack_var1) 
(pre_holding_unstack_var2) (pre_on_stack_var1_var1) (pre_on_stack_var1_var2) (pre_on_stack_var2_var1) 
(pre_on_stack_var2_var2) (pre_on_unstack_var1_var1) (pre_on_unstack_var1_var2) (pre_on_unstack_var2_var1) 
(pre_on_unstack_var2_var2) (pre_ontable_pick-up_var1) (pre_ontable_put-down_var1) (pre_ontable_stack_var1) 
(pre_ontable_stack_var2) (pre_ontable_unstack_var1) (pre_ontable_unstack_var2) (test1) (test2) (test3))

 (:action pick-up
   :parameters (?o1 - object)
   :precondition (and (not (modeProg ))(or (not (pre_ontable_pick-up_var1 ))(ontable ?o1))
(or (not (pre_clear_pick-up_var1 ))(clear ?o1))(or (not (pre_handempty_pick-up ))(handempty ))
(or (not (pre_holding_pick-up_var1 ))(holding ?o1)))
   :effect (and (when (and (del_ontable_pick-up_var1 ))(not (ontable ?o1)))
(when (and (add_ontable_pick-up_var1 ))(ontable ?o1))(when (and (del_clear_pick-up_var1 ))(not (clear ?o1)))
(when (and (add_clear_pick-up_var1 ))(clear ?o1))(when (and (del_handempty_pick-up ))(not (handempty )))
(when (and (add_handempty_pick-up ))(handempty ))(when (and (del_holding_pick-up_var1 ))(not (holding ?o1)))
(when (and (add_holding_pick-up_var1 ))(holding ?o1))))

 (:action put-down
   :parameters (?o1 - object)
   :precondition (and (not (modeProg ))(or (not (pre_ontable_put-down_var1 ))(ontable ?o1))
(or (not (pre_clear_put-down_var1 ))(clear ?o1))(or (not (pre_handempty_put-down ))(handempty ))
(or (not (pre_holding_put-down_var1 ))(holding ?o1)))
   :effect (and (when (and (del_ontable_put-down_var1 ))(not (ontable ?o1)))
(when (and (add_ontable_put-down_var1 ))(ontable ?o1))(when (and (del_clear_put-down_var1 ))(not (clear ?o1)))
(when (and (add_clear_put-down_var1 ))(clear ?o1))(when (and (del_handempty_put-down ))(not (handempty )))
(when (and (add_handempty_put-down ))(handempty ))(when (and (del_holding_put-down_var1 ))(not (holding ?o1)))
(when (and (add_holding_put-down_var1 ))(holding ?o1))))

 (:action stack
   :parameters (?o1 - object ?o2 - object)
   :precondition (and (not (modeProg ))(or (not (pre_on_stack_var1_var1 ))(on ?o1 ?o1))
(or (not (pre_on_stack_var1_var2 ))(on ?o1 ?o2))(or (not (pre_on_stack_var2_var1 ))(on ?o2 ?o1))
(or (not (pre_on_stack_var2_var2 ))(on ?o2 ?o2))(or (not (pre_ontable_stack_var1 ))(ontable ?o1))
(or (not (pre_ontable_stack_var2 ))(ontable ?o2))(or (not (pre_clear_stack_var1 ))(clear ?o1))
(or (not (pre_clear_stack_var2 ))(clear ?o2))(or (not (pre_handempty_stack ))(handempty ))
(or (not (pre_holding_stack_var1 ))(holding ?o1))(or (not (pre_holding_stack_var2 ))(holding ?o2)))
   :effect (and (when (and (del_on_stack_var1_var1 ))(not (on ?o1 ?o1)))(when (and (add_on_stack_var1_var1 ))(on ?o1 ?o1))
(when (and (del_on_stack_var1_var2 ))(not (on ?o1 ?o2)))(when (and (add_on_stack_var1_var2 ))(on ?o1 ?o2))
(when (and (del_on_stack_var2_var1 ))(not (on ?o2 ?o1)))(when (and (add_on_stack_var2_var1 ))(on ?o2 ?o1))
(when (and (del_on_stack_var2_var2 ))(not (on ?o2 ?o2)))(when (and (add_on_stack_var2_var2 ))(on ?o2 ?o2))
(when (and (del_ontable_stack_var1 ))(not (ontable ?o1)))(when (and (add_ontable_stack_var1 ))(ontable ?o1))
(when (and (del_ontable_stack_var2 ))(not (ontable ?o2)))(when (and (add_ontable_stack_var2 ))(ontable ?o2))
(when (and (del_clear_stack_var1 ))(not (clear ?o1)))(when (and (add_clear_stack_var1 ))(clear ?o1))
(when (and (del_clear_stack_var2 ))(not (clear ?o2)))(when (and (add_clear_stack_var2 ))(clear ?o2))
(when (and (del_handempty_stack ))(not (handempty )))(when (and (add_handempty_stack ))(handempty ))
(when (and (del_holding_stack_var1 ))(not (holding ?o1)))(when (and (add_holding_stack_var1 ))(holding ?o1))
(when (and (del_holding_stack_var2 ))(not (holding ?o2)))(when (and (add_holding_stack_var2 ))(holding ?o2))))

 (:action unstack
   :parameters (?o1 - object ?o2 - object)
   :precondition (and (not (modeProg ))(or (not (pre_on_unstack_var1_var1 ))(on ?o1 ?o1))
(or (not (pre_on_unstack_var1_var2 ))(on ?o1 ?o2))(or (not (pre_on_unstack_var2_var1 ))(on ?o2 ?o1))
(or (not (pre_on_unstack_var2_var2 ))(on ?o2 ?o2))(or (not (pre_ontable_unstack_var1 ))(ontable ?o1))
(or (not (pre_ontable_unstack_var2 ))(ontable ?o2))(or (not (pre_clear_unstack_var1 ))(clear ?o1))
(or (not (pre_clear_unstack_var2 ))(clear ?o2))(or (not (pre_handempty_unstack ))(handempty ))
(or (not (pre_holding_unstack_var1 ))(holding ?o1))(or (not (pre_holding_unstack_var2 ))(holding ?o2)))
   :effect (and (when (and (del_on_unstack_var1_var1 ))(not (on ?o1 ?o1)))
(when (and (add_on_unstack_var1_var1 ))(on ?o1 ?o1))(when (and (del_on_unstack_var1_var2 ))(not (on ?o1 ?o2)))
(when (and (add_on_unstack_var1_var2 ))(on ?o1 ?o2))(when (and (del_on_unstack_var2_var1 ))(not (on ?o2 ?o1)))
(when (and (add_on_unstack_var2_var1 ))(on ?o2 ?o1))(when (and (del_on_unstack_var2_var2 ))(not (on ?o2 ?o2)))
(when (and (add_on_unstack_var2_var2 ))(on ?o2 ?o2))(when (and (del_ontable_unstack_var1 ))(not (ontable ?o1)))
(when (and (add_ontable_unstack_var1 ))(ontable ?o1))(when (and (del_ontable_unstack_var2 ))(not (ontable ?o2)))
(when (and (add_ontable_unstack_var2 ))(ontable ?o2))(when (and (del_clear_unstack_var1 ))(not (clear ?o1)))
(when (and (add_clear_unstack_var1 ))(clear ?o1))(when (and (del_clear_unstack_var2 ))(not (clear ?o2)))
(when (and (add_clear_unstack_var2 ))(clear ?o2))(when (and (del_handempty_unstack ))(not (handempty )))
(when (and (add_handempty_unstack ))(handempty ))(when (and (del_holding_unstack_var1 ))(not (holding ?o1)))
(when (and (add_holding_unstack_var1 ))(holding ?o1))(when (and (del_holding_unstack_var2 ))(not (holding ?o2)))
(when (and (add_holding_unstack_var2 ))(holding ?o2))))

 (:action program_pre_ontable_pick-up_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_pick-up_var1 ))(not (del_ontable_pick-up_var1 ))
  (not (add_ontable_pick-up_var1 )))
   :effect (and (pre_ontable_pick-up_var1 )))

 (:action program_eff_ontable_pick-up_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_pick-up_var1 ))(not (add_ontable_pick-up_var1 )))
  :effect (and (when (pre_ontable_pick-up_var1 )(del_ontable_pick-up_var1 ))
  (when (not (pre_ontable_pick-up_var1 ))(add_ontable_pick-up_var1 ))))

 (:action program_pre_clear_pick-up_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_pick-up_var1 ))(not (del_clear_pick-up_var1 ))
  (not (add_clear_pick-up_var1 )))
   :effect (and (pre_clear_pick-up_var1 )))

 (:action program_eff_clear_pick-up_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_pick-up_var1 ))(not (add_clear_pick-up_var1 )))
  :effect (and (when (pre_clear_pick-up_var1 )(del_clear_pick-up_var1 ))
  (when (not (pre_clear_pick-up_var1 ))(add_clear_pick-up_var1 ))))

 (:action program_pre_handempty_pick-up
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_pick-up ))(not (del_handempty_pick-up ))
  (not (add_handempty_pick-up )))
   :effect (and (pre_handempty_pick-up )))

 (:action program_eff_handempty_pick-up
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_pick-up ))(not (add_handempty_pick-up )))
  :effect (and (when (pre_handempty_pick-up )(del_handempty_pick-up ))
  (when (not (pre_handempty_pick-up ))(add_handempty_pick-up ))))

 (:action program_pre_holding_pick-up_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_pick-up_var1 ))(not (del_holding_pick-up_var1 ))
  (not (add_holding_pick-up_var1 )))
   :effect (and (pre_holding_pick-up_var1 )))

 (:action program_eff_holding_pick-up_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_pick-up_var1 ))(not (add_holding_pick-up_var1 )))
  :effect (and (when (pre_holding_pick-up_var1 )(del_holding_pick-up_var1 ))
  (when (not (pre_holding_pick-up_var1 ))(add_holding_pick-up_var1 ))))

 (:action program_pre_ontable_put-down_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_put-down_var1 ))(not (del_ontable_put-down_var1 ))
  (not (add_ontable_put-down_var1 )))
   :effect (and (pre_ontable_put-down_var1 )))

 (:action program_eff_ontable_put-down_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_put-down_var1 ))(not (add_ontable_put-down_var1 )))
  :effect (and (when (pre_ontable_put-down_var1 )(del_ontable_put-down_var1 ))
  (when (not (pre_ontable_put-down_var1 ))(add_ontable_put-down_var1 ))))

 (:action program_pre_clear_put-down_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_put-down_var1 ))(not (del_clear_put-down_var1 ))
  (not (add_clear_put-down_var1 )))
   :effect (and (pre_clear_put-down_var1 )))

 (:action program_eff_clear_put-down_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_put-down_var1 ))(not (add_clear_put-down_var1 )))
  :effect (and (when (pre_clear_put-down_var1 )(del_clear_put-down_var1 ))
  (when (not (pre_clear_put-down_var1 ))(add_clear_put-down_var1 ))))

 (:action program_pre_handempty_put-down
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_put-down ))(not (del_handempty_put-down ))
  (not (add_handempty_put-down )))
   :effect (and (pre_handempty_put-down )))

 (:action program_eff_handempty_put-down
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_put-down ))(not (add_handempty_put-down )))
  :effect (and (when (pre_handempty_put-down )(del_handempty_put-down ))
  (when (not (pre_handempty_put-down ))(add_handempty_put-down ))))

 (:action program_pre_holding_put-down_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_put-down_var1 ))(not (del_holding_put-down_var1 ))
  (not (add_holding_put-down_var1 )))
   :effect (and (pre_holding_put-down_var1 )))

 (:action program_eff_holding_put-down_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_put-down_var1 ))(not (add_holding_put-down_var1 )))
  :effect (and (when (pre_holding_put-down_var1 )(del_holding_put-down_var1 ))
  (when (not (pre_holding_put-down_var1 ))(add_holding_put-down_var1 ))))

 (:action program_pre_on_stack_var1_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var1_var1 ))(not (del_on_stack_var1_var1 ))
  (not (add_on_stack_var1_var1 )))
   :effect (and (pre_on_stack_var1_var1 )))

 (:action program_eff_on_stack_var1_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var1_var1 ))(not (add_on_stack_var1_var1 )))
  :effect (and (when (pre_on_stack_var1_var1 )(del_on_stack_var1_var1 ))
  (when (not (pre_on_stack_var1_var1 ))(add_on_stack_var1_var1 ))))

 (:action program_pre_on_stack_var1_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var1_var2 ))(not (del_on_stack_var1_var2 ))
  (not (add_on_stack_var1_var2 )))
   :effect (and (pre_on_stack_var1_var2 )))

 (:action program_eff_on_stack_var1_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var1_var2 ))(not (add_on_stack_var1_var2 )))
  :effect (and (when (pre_on_stack_var1_var2 )(del_on_stack_var1_var2 ))
  (when (not (pre_on_stack_var1_var2 ))(add_on_stack_var1_var2 ))))

 (:action program_pre_on_stack_var2_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var2_var1 ))(not (del_on_stack_var2_var1 ))
  (not (add_on_stack_var2_var1 )))
   :effect (and (pre_on_stack_var2_var1 )))

 (:action program_eff_on_stack_var2_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var2_var1 ))(not (add_on_stack_var2_var1 )))
  :effect (and (when (pre_on_stack_var2_var1 )(del_on_stack_var2_var1 ))
  (when (not (pre_on_stack_var2_var1 ))(add_on_stack_var2_var1 ))))

 (:action program_pre_on_stack_var2_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var2_var2 ))(not (del_on_stack_var2_var2 ))
  (not (add_on_stack_var2_var2 )))
   :effect (and (pre_on_stack_var2_var2 )))

 (:action program_eff_on_stack_var2_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var2_var2 ))(not (add_on_stack_var2_var2 )))
  :effect (and (when (pre_on_stack_var2_var2 )(del_on_stack_var2_var2 ))
  (when (not (pre_on_stack_var2_var2 ))(add_on_stack_var2_var2 ))))

 (:action program_pre_ontable_stack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_stack_var1 ))(not (del_ontable_stack_var1 ))
  (not (add_ontable_stack_var1 )))
   :effect (and (pre_ontable_stack_var1 )))

 (:action program_eff_ontable_stack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_stack_var1 ))(not (add_ontable_stack_var1 )))
  :effect (and (when (pre_ontable_stack_var1 )(del_ontable_stack_var1 ))
  (when (not (pre_ontable_stack_var1 ))(add_ontable_stack_var1 ))))

 (:action program_pre_ontable_stack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_stack_var2 ))(not (del_ontable_stack_var2 ))
  (not (add_ontable_stack_var2 )))
   :effect (and (pre_ontable_stack_var2 )))

 (:action program_eff_ontable_stack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_stack_var2 ))(not (add_ontable_stack_var2 )))
  :effect (and (when (pre_ontable_stack_var2 )(del_ontable_stack_var2 ))
  (when (not (pre_ontable_stack_var2 ))(add_ontable_stack_var2 ))))

 (:action program_pre_clear_stack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_stack_var1 ))(not (del_clear_stack_var1 ))
  (not (add_clear_stack_var1 )))
   :effect (and (pre_clear_stack_var1 )))

 (:action program_eff_clear_stack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_stack_var1 ))(not (add_clear_stack_var1 )))
  :effect (and (when (pre_clear_stack_var1 )(del_clear_stack_var1 ))
  (when (not (pre_clear_stack_var1 ))(add_clear_stack_var1 ))))

 (:action program_pre_clear_stack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_stack_var2 ))(not (del_clear_stack_var2 ))
  (not (add_clear_stack_var2 )))
   :effect (and (pre_clear_stack_var2 )))

 (:action program_eff_clear_stack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_stack_var2 ))(not (add_clear_stack_var2 )))
  :effect (and (when (pre_clear_stack_var2 )(del_clear_stack_var2 ))
  (when (not (pre_clear_stack_var2 ))(add_clear_stack_var2 ))))

 (:action program_pre_handempty_stack
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_stack ))(not (del_handempty_stack ))
  (not (add_handempty_stack )))
   :effect (and (pre_handempty_stack )))

 (:action program_eff_handempty_stack
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_stack ))(not (add_handempty_stack )))
  :effect (and (when (pre_handempty_stack )(del_handempty_stack ))
  (when (not (pre_handempty_stack ))(add_handempty_stack ))))

 (:action program_pre_holding_stack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_stack_var1 ))(not (del_holding_stack_var1 ))
  (not (add_holding_stack_var1 )))
   :effect (and (pre_holding_stack_var1 )))

 (:action program_eff_holding_stack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_stack_var1 ))(not (add_holding_stack_var1 )))
  :effect (and (when (pre_holding_stack_var1 )(del_holding_stack_var1 ))
  (when (not (pre_holding_stack_var1 ))(add_holding_stack_var1 ))))

 (:action program_pre_holding_stack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_stack_var2 ))(not (del_holding_stack_var2 ))
  (not (add_holding_stack_var2 )))
   :effect (and (pre_holding_stack_var2 )))

 (:action program_eff_holding_stack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_stack_var2 ))(not (add_holding_stack_var2 )))
  :effect (and (when (pre_holding_stack_var2 )(del_holding_stack_var2 ))
  (when (not (pre_holding_stack_var2 ))(add_holding_stack_var2 ))))

 (:action program_pre_on_unstack_var1_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var1_var1 ))(not (del_on_unstack_var1_var1 ))
  (not (add_on_unstack_var1_var1 )))
   :effect (and (pre_on_unstack_var1_var1 )))

 (:action program_eff_on_unstack_var1_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var1_var1 ))(not (add_on_unstack_var1_var1 )))
  :effect (and (when (pre_on_unstack_var1_var1 )(del_on_unstack_var1_var1 ))
  (when (not (pre_on_unstack_var1_var1 ))(add_on_unstack_var1_var1 ))))

 (:action program_pre_on_unstack_var1_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var1_var2 ))(not (del_on_unstack_var1_var2 ))
  (not (add_on_unstack_var1_var2 )))
   :effect (and (pre_on_unstack_var1_var2 )))

 (:action program_eff_on_unstack_var1_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var1_var2 ))(not (add_on_unstack_var1_var2 )))
  :effect (and (when (pre_on_unstack_var1_var2 )(del_on_unstack_var1_var2 ))
  (when (not (pre_on_unstack_var1_var2 ))(add_on_unstack_var1_var2 ))))

 (:action program_pre_on_unstack_var2_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var2_var1 ))(not (del_on_unstack_var2_var1 ))
  (not (add_on_unstack_var2_var1 )))
   :effect (and (pre_on_unstack_var2_var1 )))

 (:action program_eff_on_unstack_var2_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var2_var1 ))(not (add_on_unstack_var2_var1 )))
  :effect (and (when (pre_on_unstack_var2_var1 )(del_on_unstack_var2_var1 ))
  (when (not (pre_on_unstack_var2_var1 ))(add_on_unstack_var2_var1 ))))

 (:action program_pre_on_unstack_var2_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var2_var2 ))(not (del_on_unstack_var2_var2 ))
  (not (add_on_unstack_var2_var2 )))
   :effect (and (pre_on_unstack_var2_var2 )))

 (:action program_eff_on_unstack_var2_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var2_var2 ))(not (add_on_unstack_var2_var2 )))
  :effect (and (when (pre_on_unstack_var2_var2 )(del_on_unstack_var2_var2 ))
  (when (not (pre_on_unstack_var2_var2 ))(add_on_unstack_var2_var2 ))))

 (:action program_pre_ontable_unstack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_unstack_var1 ))(not (del_ontable_unstack_var1 ))
  (not (add_ontable_unstack_var1 )))
   :effect (and (pre_ontable_unstack_var1 )))

 (:action program_eff_ontable_unstack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_unstack_var1 ))(not (add_ontable_unstack_var1 )))
  :effect (and (when (pre_ontable_unstack_var1 )(del_ontable_unstack_var1 ))
  (when (not (pre_ontable_unstack_var1 ))(add_ontable_unstack_var1 ))))

 (:action program_pre_ontable_unstack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_unstack_var2 ))(not (del_ontable_unstack_var2 ))
  (not (add_ontable_unstack_var2 )))
   :effect (and (pre_ontable_unstack_var2 )))

 (:action program_eff_ontable_unstack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_unstack_var2 ))(not (add_ontable_unstack_var2 )))
  :effect (and (when (pre_ontable_unstack_var2 )(del_ontable_unstack_var2 ))
  (when (not (pre_ontable_unstack_var2 ))(add_ontable_unstack_var2 ))))

 (:action program_pre_clear_unstack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_unstack_var1 ))(not (del_clear_unstack_var1 ))
  (not (add_clear_unstack_var1 )))
   :effect (and (pre_clear_unstack_var1 )))

 (:action program_eff_clear_unstack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_unstack_var1 ))(not (add_clear_unstack_var1 )))
  :effect (and (when (pre_clear_unstack_var1 )(del_clear_unstack_var1 ))
  (when (not (pre_clear_unstack_var1 ))(add_clear_unstack_var1 ))))

 (:action program_pre_clear_unstack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_unstack_var2 ))(not (del_clear_unstack_var2 ))
  (not (add_clear_unstack_var2 )))
   :effect (and (pre_clear_unstack_var2 )))

 (:action program_eff_clear_unstack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_unstack_var2 ))(not (add_clear_unstack_var2 )))
  :effect (and (when (pre_clear_unstack_var2 )(del_clear_unstack_var2 ))
  (when (not (pre_clear_unstack_var2 ))(add_clear_unstack_var2 ))))

 (:action program_pre_handempty_unstack
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_unstack ))(not (del_handempty_unstack ))
  (not (add_handempty_unstack )))
   :effect (and (pre_handempty_unstack )))

 (:action program_eff_handempty_unstack
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_unstack ))(not (add_handempty_unstack )))
  :effect (and (when (pre_handempty_unstack )(del_handempty_unstack ))
  (when (not (pre_handempty_unstack ))(add_handempty_unstack ))))

 (:action program_pre_holding_unstack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_unstack_var1 ))(not (del_holding_unstack_var1 ))
  (not (add_holding_unstack_var1 )))
   :effect (and (pre_holding_unstack_var1 )))

 (:action program_eff_holding_unstack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_unstack_var1 ))(not (add_holding_unstack_var1 )))
  :effect (and (when (pre_holding_unstack_var1 )(del_holding_unstack_var1 ))
  (when (not (pre_holding_unstack_var1 ))(add_holding_unstack_var1 ))))

 (:action program_pre_holding_unstack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_unstack_var2 ))(not (del_holding_unstack_var2 ))
  (not (add_holding_unstack_var2 )))
   :effect (and (pre_holding_unstack_var2 )))

 (:action program_eff_holding_unstack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_unstack_var2 ))(not (add_holding_unstack_var2 )))
  :effect (and (when (pre_holding_unstack_var2 )(del_holding_unstack_var2 ))
  (when (not (pre_holding_unstack_var2 ))(add_holding_unstack_var2 ))))

 (:action validate_1
   :parameters ()
   :precondition (and (modeProg ))
  :effect (and (not (modeProg ))(clear a)(not (clear b))(clear c)(not (clear d))(not (clear e))(not (clear f))
  (not (clear g))(handempty )(not (holding a))(not (holding b))(not (holding c))(not (holding d))
  (not (holding e))(not (holding f))(not (holding g))(not (on a a))(not (on a b))(not (on a c))(not (on a d))
  (not (on a e))(not (on a f))(on a g)(not (on b a))(not (on b b))(not (on b c))(not (on b d))(on b e)
  (not (on b f))(not (on b g))(not (on c a))(not (on c b))(not (on c c))(on c d)(not (on c e))(not (on c f))
  (not (on c g))(not (on d a))(on d b)(not (on d c))(not (on d d))(not (on d e))(not (on d f))(not (on d g))
  (not (on e a))(not (on e b))(not (on e c))(not (on e d))(not (on e e))(on e f)(not (on e g))(not (on f a))
  (not (on f b))(not (on f c))(not (on f d))(not (on f e))(not (on f f))(not (on f g))(not (on g a))
  (not (on g b))(not (on g c))(not (on g d))(not (on g e))(not (on g f))(not (on g g))(not (ontable a))
  (not (ontable b))(not (ontable c))(not (ontable d))(not (ontable e))(ontable f)(ontable g)(test1 )))

 (:action validate_2
   :parameters ()
  :precondition (and (not (modeProg ))(clear a)(clear b)(clear c)(clear d)(clear e)(not (clear f))(not (clear g))
  (handempty )(not (holding a))(not (holding b))(not (holding c))(not (holding d))(not (holding e))
  (not (holding f))(not (holding g))(not (on a a))(not (on a b))(not (on a c))(not (on a d))(not (on a e))
  (not (on a f))(on a g)(not (on b a))(not (on b b))(not (on b c))(not (on b d))(not (on b e))(not (on b f))
  (not (on b g))(not (on c a))(not (on c b))(not (on c c))(not (on c d))(not (on c e))(not (on c f))(not (on c g))
  (not (on d a))(not (on d b))(not (on d c))(not (on d d))(not (on d e))(not (on d f))(not (on d g))(not (on e a))
  (not (on e b))(not (on e c))(not (on e d))(not (on e e))(on e f)(not (on e g))(not (on f a))(not (on f b))
  (not (on f c))(not (on f d))(not (on f e))(not (on f f))(not (on f g))(not (on g a))(not (on g b))(not (on g c))
  (not (on g d))(not (on g e))(not (on g f))(not (on g g))(not (ontable a))(ontable b)(ontable c)(ontable d)
  (not (ontable e))(ontable f)(ontable g)(test1 ))
   :effect (and (test2 )(not (test1 ))))

 (:action validate_3
   :parameters ()
  :precondition (and (not (modeProg ))(clear a)(clear b)(clear c)(clear d)(clear e)(clear f)(clear g)(handempty )
  (not (holding a))(not (holding b))(not (holding c))(not (holding d))(not (holding e))(not (holding f))
  (not (holding g))(not (on a a))(not (on a b))(not (on a c))(not (on a d))(not (on a e))(not (on a f))
  (not (on a g))(not (on b a))(not (on b b))(not (on b c))(not (on b d))(not (on b e))(not (on b f))(not (on b g))
  (not (on c a))(not (on c b))(not (on c c))(not (on c d))(not (on c e))(not (on c f))(not (on c g))(not (on d a))
  (not (on d b))(not (on d c))(not (on d d))(not (on d e))(not (on d f))(not (on d g))(not (on e a))(not (on e b))
  (not (on e c))(not (on e d))(not (on e e))(not (on e f))(not (on e g))(not (on f a))(not (on f b))(not (on f c))
  (not (on f d))(not (on f e))(not (on f f))(not (on f g))(not (on g a))(not (on g b))(not (on g c))(not (on g d))
  (not (on g e))(not (on g f))(not (on g g))(ontable a)(ontable b)(ontable c)(ontable d)(ontable e)(ontable f)
  (ontable g)(test2 ))
   :effect (and (not (test2 ))(test3 )))

)

  \end{verbatim}
\end{scriptsize}

\begin{figure}[hbtp!]
\begin{scriptsize}  
  \begin{verbatim}
(define (problem learning_problem)
  (:domain blocks)
  (:objects g -  object c -  object d -  object e -  object a -  object b -  object f -  object )
  (:init (modeProg ) )
  (:goal (and (test3 ))))
  \end{verbatim}
\end{scriptsize}  
\caption{\small Compiled PDDL problem file for learning the blocksworld action models from two initial and final states.}
\label{fig:compiled-problem}
\end{figure}